### 2016-11-22 intro.tex
Made Hugh's suggested changes

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$ 
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b \sim 0.05$), which existed as a nearly isothermal
($\delta T/T\sim10^{-5}$) gas of hydrogen and helium at the earliest
observational time.  All of this began roughly $13.8$ billion years ago, with
the universe emerging from a state of high temperature and density,
expanding to form what we see today.  The primordial elemental abundances, with
$\sim4/5$ of the universe's baryons in hydrogen, and the remainder in helium,
was established just a few minutes after the big bang, during a brief period of
``Big Bang Nucleosynthesis'' \citep{Alpher1948}.  The remaining heavier elements
have been formed via stellar nucleosynthesis, through fusion in the cores of
stars and in energetic processes that take place during supernovae (SN)
\citep{Wagoner1967}, or through interactions between cosmic rays and nucleii in
the interstellar medium (ISM) \citep{Reeves1970}.

Much of what we know about the cosmology of our universe comes from
observations of the first light to escape the universe's surface of last
scattering as it became optically thin: the cosmic microwave background
radiation (CMB).  Prior to a redshift of $z\approx1100$, the universe had a
high ionization fraction such that the mean free path for photons was short,
making all of space effectively opaque.  Once the universe became cool enough
that hydrogen became fully neutral, the optical depth fell low enough that
photons could free-stream, decoupled from baryons.  This ultraviolet light we
observe today, redshifted to microwaves with a temperature of
$\sim2.7\;\rm{K}$.  When the power spectrum of this radiation is measured
\citep{Spergel2003}, features in that spectrum can be used to determine the
components of the universe (with some ambiguity that can be resolved via SN Ia
and large scale structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}), as well as the
spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation
\citep{Guth1981,Linde1982}.  These perturbations grow through
gravitational collapse.  Until $z\sim1100$, only dark matter perturbations were
able to grow.  Prior to this, baryons were coupled to photons, causing density
perturbations to oscillate as stable sound waves.  This early collapse
of matter meant that when decoupling occurred at $z\sim1100$, baryons were able
to collapse into the deeper potential wells grown out of dark-matter density
perturbations.  These overdensities expand slower that the surrounding medium,
as they experience a drag from their own self-gravity.  Eventually, these
regions stop expanding altogether and begin collapsing at the turnaround time
$t_{turn}$, and virialize at $2\;t_{turn}$ \citep{Peebles1980}. The first
galaxies likely began to form around $z\sim20-50$.

@@

The formation of massive stars takes place mostly in clusters of $10^4$ or more
stars.  This means that the SN of these stars can be treated not as
individual dumps of $10^{51}\;\rm{erg/SN}$, but instead as a luminosity of
$10^{34}\;\rm{\frac{erg}{s \cdot M_\odot}}$.  This allows us to use stellar wind
models like \citet{Weaver1977} to study the evolution of superbubbles driven by
clustered SN.  Each individual SN's ejecta thermalizes in a small
region within the center of the bubble, and the resulting hot gas drives a shock
that sweeps up the surrounding ISM.  In a short time, the swept-up ISM can
radiatively cool efficiently , resulting in a thin, cold shell surrounding a
hot, diffuse bubble.

>>>>

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b \sim 0.05$), which existed as a nearly isothermal
($\delta T/T\sim10^{-5}$) gas of hydrogen and helium at the earliest
observational time.  The composition of the universe, and the precise value of
the mass-energy density for these components sets the lifetime of the universe,
how rapidly it expands (or contracts), and has tremendous impact on the growth
of structure.  This structure is ultimately where galaxies will form. All of
this began $13.8$ billion years ago, with the universe emerging from a state of
high temperature and density, expanding to form what we see today.  The
primordial elemental abundances, with $\sim4/5$ of the universe's baryons in
hydrogen, and the remainder in helium, was established just a few minutes after
the big bang, during a brief period of ``Big Bang Nucleosynthesis''
\citep{Alpher1948}.  The remaining heavier elements have been formed via stellar
nucleosynthesis, through fusion in the cores of stars and in energetic processes
that take place during supernovae (SN) \citep{Wagoner1967}, or through
interactions between cosmic rays and nucleii in the interstellar medium (ISM)
\citep{Reeves1970}.  The chemical composition of the universe establishes the
kinds of stars that form within galaxies and the rates at which gas within the
ISM and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  This ultraviolet light we observe today, redshifted to
microwaves with a temperature of $\sim2.7\;\rm{K}$.  When the power spectrum of
this radiation is measured \citep{Spergel2003}, features in that spectrum can be
used to determine the components of the universe (with some ambiguity that can
be resolved via SN Ia and large scale structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}), as well as the
spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation \citep{Guth1981,Linde1982}.  These
perturbations grow through gravitational collapse.  Until $z\sim1100$, only dark
matter perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark-matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding altogether and begin collapsing at the
turnaround time $t_{turn}$, and virialize at $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$.  Without the
earlier growth of dark matter perturbations, structure formation would have
occured much later, and the number of galaxies and distribution of their masses
would be different than those we see today \citep{Davis1985}.

@@

\subsection{Superbubbles: The Natural Scale for Feedback}
The formation of massive stars takes place mostly in clusters of $10^4$ or more
stars.  This means that the SN of these stars can be treated not as
individual dumps of $10^{51}\;\rm{erg/SN}$, but instead as a luminosity of
$10^{34}\;\rm{\frac{erg}{s \cdot M_\odot}}$.  This allows us to use stellar wind
models like \citet{Weaver1977} to study the evolution of superbubbles driven by
clustered SN.  Each individual SN's ejecta thermalizes in a small
region within the center of the bubble, and the resulting hot gas drives a shock
that sweeps up the surrounding ISM.  In a short time, the swept-up ISM can
radiatively cool efficiently , resulting in a thin, cold shell surrounding a
hot, diffuse bubble.

A wealth of studies have
